Software

### 2025-05-02 about.qmd
Full update

---
title: "Members"
---

[Marianne Huebner](https://cstat.msu.edu/people/marianne-huebner-phd-sheher) (chair), Professor of Statistics and Probability at Michigan State University, USA. <https://orcid.org/0000-0002-9694-9231>

Carsten Oliver Schmidt (chair), Professor of the Institute of Community Medicine, University Medicine of Greifswald

Mark Baillie, Advanced Methodology and Data Science, Clinical Development and Analytics. Novartis Pharma AG, Basel, Switzerland

Saskia le Cessie, Professor at Leiden University Medical Center, Netherlands

[Lara Lusa](https://www.famnit.upr.si/sl/zaposleni-in-sodelavci/lara.lusa/), Professor at the Faculty of Mathematics, Natural Sciences and Information Technologies of the University of Primorska in Koper, Slovenia

[Michael Schell](https://www.moffitt.org/research-science/researchers/michael-schell/), Moffit Cancer Center, Florida, USA

## **Bios**

**Mark Baillie** is a statistician supporting the clinical development and analytics department at Novartis. He is a member of the methodology and consulting group, providing strategic and methodological input to various clinical and non-clinical priority projects across the company. He has a focus on effective visual communication working on a number of internal and external initiatives to improve the reporting of clinical trials and exploratory projects.

**Marianne Huebner** is Professor of the Department of Statistics and Probability and Director of the Center for Statistical Training and Consulting at Michigan State University. After receiving a PhD in Applied Mathematics she held positions at UC Berkeley, the Mathematical Sciences Research Institute in Berkeley, the Mayo Clinic, and the University Medical Center in Hamburg, Germany. She develops and applies statistical models motivated by scientific questions. Her research interests include modeling health outcomes, initial data analyses, and sports and exercise.

**Saskia le Cessie** is a statistician with a joint appointment at the Department of Clinical Epidemiology and the Section Medical Statistics, Department of Biomedical Datasciences. She obtained a master in Mathematics (with minor in Computer Sciences) at the University of Utrecht and obtained her PhD in Medical Statistics at the University of Leiden. She is a broadly oriented statistician and involved in several large epidemiological studies of the Department of Clinical Epidemiology. Her research interests are in epidemiological and statistical methods for observational studies, in particular instrumental variable analysis, mediation analysis, competing risks, causal modelling, meta analysis and repeated measurements.

**Lara Lusa** is Professor of Statistics at the Faculty of Mathematics, Natural Sciences and Information Technologies of the University of Primorska, Slovenia. After receiving a PhD in Applied Statistics from the University of Florence, she worked at the Italian National Cancer Institute of Milan, at the Institute of Biostatistics and Medical Informatics of the University of Ljubljana and was a visiting fellow at the Biometric Research Branch of the National Cancer Institute in Bethesda. Her current research interests are related to the development and validation of predictive models in bio‐medicine, with particular attention to applications using high‐dimensional data and rare events. She is also interested in statistical programming, design of simulation studies, and teaching statistics. Beside the methodological research in statistics, she is also heavily involved as a statistical consultant to researchers in medicine.

**Michael Schell**

**Carsten Oliver Schmidt** is Professor at the Institute of Community Medicine, University Medicine of Greifswald. He is deputy head of the Department of SHIP/ Clinical-Epidemiological Research and leads the functional division "Quality in the Health Sciences". He is PI of several methods and research quality related projects, and speaker of the methods section of the German Society for Epidemiology. He is responsible for quality management procedures in population based observational health studies, has expertise in data linkage of primary and secondary data including registries, claims data, hospitals, and in the development of research infrastructures. Research topics include research methods, the study of subclinical and clinical disorders, incidental findings.

>>>>

---
title: "About this website"
---

**Andreas Klinger**: set-up and maintenance

**Georg Heinze**: supervision and revision

Last changes: 2 May 2025.

### Acknowledgements

Many thanks to Marianne Huebner, who designed the website of the STRATOS Topic Group 3, which the website of Topic Group 2 is based on.

### Address for correspondence

STRATOS TG2 c/o Institute of Clinical Biometrics, center for Medical Data Science, Medical University of Vienna, Vienna, Austria. E-mail: stratos.tg2@gmail.com.

### 2025-05-02 activities.qmd
Full update

NOTE: This is the text that was originally on the TG2 website and it is severely outdated!

* Georg Heinze and Christine Wallisch will give a STRATOS-TG2 workshop on 'Variable selection – a review and recommendations for the practicing statistician' at the IBS-ROeS conference in Salzburg, 6 September 2021

* Aris Perperoglou will host a session on "STRATOS initiative - on various issues in multivariable model building" at the RSS International Conference in Manchester,  6-9 September, with talks from Bianca de Stavola, Tim Morris, Willi Sauerbrei and Michal Abrahamowicz.

* Georg Heinze and Mark Baillie will host a session on 'STRATOS initiative - Tools for guidance for analysis of observational studies' at the IBS-ROeS conference in Salzburg, 7-10 September 2021, featuring Rolf Groenwold from TG2 with a talk on ' Using videos and other media for guidance'

>>>>

## Recent

* Georg Heinze and Theresa Ullmann presented a 3-hour Statistics in Practice Showcase session at the International Biometric Conference in Atlanta, GA, USA, December 2024. Find out more at [https://www.ibc2024.org](https://www.ibc2024.org/events/showcase). [The slides are available at OSF here](https://osf.io/a5gtk/).

## Upcoming

* Internal TG2 meeting in Vienna, 7-9 May 2025 at the [Center for Medical Data Science](https://data-science.meduniwien.ac.at), including a public talk by Matthias Schmid on [Achieving explainable machine learning by functional decomposition of black-box models into explainable predictor effects](https://viennabiometrics.github.io/abstracts/20250805_WBS_Schmid.pdf)

* Minisymposium at the International Society of Clinical Biostatistics (ISCB) conference in Basel, CH, 28 Aug 2025. Topic of the session: Statistical Research needs to improve – on the important roles of simulation studies and guidance for analysis. Speakers (among others): Georg Heinze, Willi Sauerbrei, Michal Abrahamowicz [Find out more](https://iscb2025.info/mini-symposia.html#mini-symposium-3-statistical-research-needs-to-improve-on-the-important-roles-of-simulation-studies-and-guidance-for-analysis)

* Invited session at the [ROeS 2025](https://roes2025.at/index.html) (Conference of the Austro-Swiss Region of the International Biometric Society) in Graz, A, 14-18 Sep 2025. Topic of the session: Contrasting Bayesian and frequentist model building for descriptive research questions - a paired-design experiment of STRATOS-TG2. Invited speaker: Mariana Nold, further speakers for TG2: Michael Kammer, Georg Heinze, Aliaksandr Hubin

* Topic-contributed paper session at Joint Statistical Meeting (JSM), 6 Aug 2025. Speaker for TG2: Marc Henrion. Topic of the talk: STRATOS Topic Group 2 – Selection of variables and functional forms in multivariable analysis: Results from a review of Covid-19 prognostic models.

### 2025-05-02 members.qmd
Full update

[Georg Heinze](https://www.meduniwien.ac.at/researcher/georg_heinze) (chair): Institute of Clinical Biometrics, Center for Medical Statistics, Informatics and Intelligent Systems, Medical University of Vienna,  Austria.

@@

**Heiko Becher**: Institute for Medical Biometry and Epidemiology, University Medical Center Hamburg-Eppendorf, Hamburg, Germany

@@

**Daniela Dunkler**: Institute of Clinical Biometrics, Center for Medical Statistics, Informatics and Intelligent Systems, Medical University of Vienna,  Austria

**Rolf Groenwold**:  Leiden University Medical Centre, Leiden, The Netherlands

**Frank Harrell**: School of Medicine, Vanderbilt University Medical Center, USA

**Nadja Klein**: School of Business and Economics , Humboldt-University of Berlin, Germany

**Geraldine Rauch**:  Institute of Biometry and Clinical  Epidemiology, Charité – Universitätsmedizin Berlin , Germany

**Patrick Royston**: Department of Statistical Science, UCL, UK

**Matthias Schmid**:  Institute for Medicine Biometry, Informatics and Epidemiology, Bonn, Germany

@@

**Michael Kammer**: Institute of Clinical Biometrics, Center for Medical Statistics, Informatics and Intelligent Systems, Medical University of Vienna,  Austria

@@

**Kim Luijken**:  Leiden University Medical Centre, Leiden, The Netherlands

**Christine Wallisch**: Institute of Clinical Biometrics, Center for Medical Statistics, Informatics and Intelligent Systems, Medical University of Vienna,  Austria

>>>>

**[Georg Heinze](https://www.meduniwien.ac.at/researcher/georg_heinze) (chair)**: Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria.

@@

@@

**Daniela Dunkler**: Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria

**Frank Harrell**: School of Medicine, Vanderbilt University Medical Center, Nashville, TN, USA

**Marc Henrion**: Malawi Liverpool Wellcome Research Programme, Blantyre, Malawi and Department of Clinical Sciences, Liverpool School of Tropical Medicine, Liverpool, UK

**Michael Kammer**: Division of Nephrology, Department of Medicine III and Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria

**Mariana Nold**: Department of Sociology, Friedrich-Schiller University of Jena, Jena, Germany

**Matthias Schmid**:  Institute for Medicine Biometry, Informatics and Epidemiology, University of Bonn, Bonn, Germany

@@

**Gregor Buch**: Preventive Cardiology and Preventive Medicine, Department of Cardiology, University Medical Center of the Johannes Gutenberg University Mainz, Mainz, Germany

@@

**Theresa Ullmann**: Institute of Clinical Biometrics, Center for Medical Data Science, Medical University of Vienna,  Austria

### 2025-05-02 resources.qmd
Full update

## R packages on CRAN

* Regression Modeling Strategies: [rms](https://cran.r-project.org/web/packages/rms/index.html) Regression modeling, testing, estimation, validation, graphics, prediction, and typesetting by storing enhanced model design attributes in the fit. 'rms' is a collection of functions that assist with and streamline modeling.

* Augmented Backward Elimination: [abe](https://cran.r-project.org/web/packages/abe/index.html) Performs augmented backward elimination and checks the stability of the obtained model. Augmented backward elimination combines significance or information based criteria with the change in estimate to either select the optimal model for prediction purposes or to serve as a tool to obtain a practically sound, highly interpretable model. More details can be found in [Dunkler et al. (2014)](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0113677).

* Multivariable Fractional Polynomials: [mfp](https://cran.r-project.org/web/packages/mfp/index.html)  Fractional polynomials are used to represent curvature in regression models. Software page: [https://mfp.imbi.uni-freiburg.de/software](https://mfp.imbi.uni-freiburg.de/software)

>>>>

## R packages and other ressources provided by TG2 members

### Regression Modeling Strategies

Regression modeling, testing, estimation, validation, graphics, prediction, and typesetting by storing enhanced model design attributes in the fit. 'rms' is a collection of functions that assist with and streamline modeling.

* R package: [rms](https://cran.r-project.org/web/packages/rms/index.html) 
* Online book companion: [https://hbiostat.org/doc/rms/book/](https://hbiostat.org/doc/rms/book/)

### Multivariable Fractional Polynomials (MFP)

MFP is a pragmatic procedure to create a multivariable model with the twin aims of selecting important variables and determining a suitable functional form for continuous predictors.

Ressources:

* R-package [mfp](https://cran.r-project.org/web/packages/mfp/index.html)
* R-package [mfp2](https://cran.r-project.org/package=mfp2)
* MFP website with explanations, examples, software, links, etc.: [https://mfp.imbi.uni-freiburg.de/software](https://mfp.imbi.uni-freiburg.de/software)

### Augmented Backward Elimination

Augmented backward elimination combines significance or information based criteria with the change in estimate to either select the optimal model for prediction purposes or to serve as a tool to obtain a practically sound, highly interpretable model. More details can be found in [Dunkler et al. (2014)](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0113677).

* R package: [abe](https://cran.r-project.org/web/packages/abe/index.html) Performs augmented backward elimination and checks 
the stability of the obtained model.

### Tetris-plot

The Tetris-plot is a tool to visualize the instability of models obtained by variable selection.

* Github website with R package (development version): https://gsk-biostatistics.github.io/tetrisplot
